|

### 2020-05-07 paper/Manuscript.Rmd
Fix tests and synthetic

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Loek Brinkman"
    affiliation   : "2,3"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Open Science Community Utrecht"
  - id            : "3"
    institution   : "UMC Utrecht"

@@

The social sciences are amidst a paradigm shift towards open science. Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [e.g., @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018]. However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012]. Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012]. Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].

@@

A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of open science. <!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019. Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5. Peikert and Brandmaier aim to achieve strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and reinstated on different systems, instead of on a private computer. The workflow is excellent, but difficult to implement for non-programmers. WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*. This should be sufficient for most users, but can also be combined with Peikert and Brandmaier's approach if strict computational reproducibility is essential. WORCS thus addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines. Still, it can be combined with existing approaches.

@@

WORCS relies on a few free, open source software solutions that, when combined, facilitate meeting the TOP-guidelines related to citation, sharing, and preregistration. The first tool is *dynamic document generation*. Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses. The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables. Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text. Although there is a slight learning curve to transitioning to DDG, the investment really pays off:

@@

The second tool is *version control*: Maintaining an indelible log of every change to all project files. Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016]. Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions. WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit". Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results). A project version controlled with Git is called a "Repository", or Repo.

@@

The third tool is *dependency management*: Keeping track of exactly what software was used to conduct the analyses. <!-- Being able to replicate analyses requires not only access to the data and analysis code, but also to the software used to evaluate that code.--> At first glance, it might seem sufficient to state that analyses were conducted in *Program X*. However, every program is susceptible to changes and updates. R-packages in particular are updated frequently, because they are Open Source, and there is an active community of developers contributing functionality and bugfixes. Potentially, any update could change the results of the code. For example, R-version 3.6.0 fixed a bug in the random number generator. This means that any analyses that use random numbers, such as bootstrapping, cannot be strictly replicated from previous versions.

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine. This is quite effective, but very difficult to set up. <!-- Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.

>>>>

author: 
  - name          : "Caspar J. van Lissa"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email         : "c.j.vanlissa@uu.nl"
  - name          : "Andreas M. Brandmaier"
    affiliation   : "3,4"
  - name          : "Loek Brinkman"
    affiliation   : "2,5"

affiliation:
  - id            : "1"
    institution   : "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id            : "2"
    institution   : "Open Science Community Utrecht"
  - id            : "3"
    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id            : "4"
    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id            : "5"
    institution   : "UMC Utrecht"

@@

The social sciences are amidst a paradigm shift towards open science. Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018]. However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012]. Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012]. Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].

@@

A different class of solutions instead focuses on the practical issue of *how* researchers can meet the requirements of open science. <!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example here is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019. Being limited to computational reproducibility, this workflow only addresses TOP-guidelines 2, 3, and 5. Peikert and Brandmaier's approach overlaps, to some extent, with the solution proposed here, but focuses on achieving strict computational reproducibility by conducting analyses on a virtual machine, that can be stored like a sort of time capsule and identically reinstated on different systems. The workflow is excellent, but difficult to implement for non-programmers. WORCS instead uses a much more lightweight approach to safeguard computational reproducibility under *most circumstances*. This should be sufficient for most users, but can also be combined with Peikert and Brandmaier's approach if strict computational reproducibility is essential. WORCS addresses a unique issue not covered by existing solutions, namely to provide a workflow most conducive to satisfying the TOP-guidelines, while being compatible with existing approaches.

@@

WORCS relies on several free, open source software solutions. The first of these is *dynamic document generation*. Dynamic document generation (DDG) means writing scientific reports in a format that interleaves written reports with blocks of computer code used to conduct the analyses. The text is automatically formatted as an APA-style manuscript [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020], and results of the code blocks are automatically generated and insterted in the text, or rendered as Figures and Tables. Dynamic document generation supercedes the classical approach of using separate programs to write prose and conduct analyses, and then manually copy-pasting analysis results into the text. Although there is a slight learning curve to transitioning to DDG, the investment really pays off:

@@

The second solution is *version control*: Maintaining an indelible log of every change to all project files. Version control is a near-essential tool for scientific reproducibility, as anyone learns who has had the experience of accidentally deleting a crucial file, or of being unable to reproduce analyses because they ran some analyses interactively instead of using syntax [see also @blischakQuickIntroductionVersion2016]. Many scientists use some form of *implicit* version control; for example, by renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"), tracking changes in word processing software, or using cloud hosting services that retain backups of previous versions. WORCS instead uses the *explicit* version control software "Git". Git tracks changes to files, and stores these changes when the user makes a "commit". Git retains a complete log of all commits, and users can compare changes between different commits, or go back to a previous version of the code (for example, after making a mistake, or to replicate a previous version of the results). A project version controlled with Git is called a "Repository", or Repo.

@@

The third solution is *dependency management*: Keeping track of exactly what software was used to conduct the analyses. <!-- Being able to replicate analyses requires not only access to the data and analysis code, but also to the software used to evaluate that code.--> At first glance, it might seem sufficient to state that analyses were conducted in *Program X*. However, every program is susceptible to changes and updates. R-packages in particular are updated frequently, because they are Open Source, and there is an active community of developers contributing functionality and bugfixes. Potentially, any update could change the results of the code. For example, R-version 3.6.0 fixed a bug in the random number generator. This means that any analyses that use random numbers, such as bootstrapping, cannot be strictly replicated from previous versions. Or, for example, R-version 4.0.0 changed a default parameter that affects how data are loaded from files.

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs. These solutions differ in user-friendliness and effectiveness. If strict reproducibility is required, one might use the aforementioned Docker-based workflow by @peikertReproducibleDataAnalysis2019, which runs all analyses inside a virtual machine. This is effective, and also pertains to software outside the R environment (e.g., preprocessing tools for neuroimaging data). However, it is very difficult to set up. <!--For small projects, or novice users || Mention CodeOcean here-->A slightly less sophisticated solution is offered by the R-package `packrat`, which installs all software used in a project to a dedicated library that can be stored and copied with the project. This solution is reasonably user-friendly and effective, but has the disadvantage that for each project, all packages must be installed anew, which can take a long time (sometimes hours) and a lot of hard drive space. Moreover, Git/GitHub is not designed to store your entire project library, so `packrat` does not address the issue of how other users are supposed to get access to your library. In conclusion, `packrat` is somewhat unwieldy. The package `checkpoint` offers an extremely lightweight solution: Instead of storing all software used, it simply records the calendar date, and then installs the version of all required packages that was available on the central R repository `CRAN` on that date from an archival website. Thus, all that is needed to restore a project is the date on which it was conducted. This is very user-friendly, and reasonably effective - the only problem is that this solution is limited to `CRAN` packages, and cannot mix package versions from different dates.
